Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-11 popl24/hyperprop.tex
checkpoint

As we've discussed in our section on Related Work
(\ref{section-related}), previous authors have noted connections
between the traditional formulation of the MPC security model- aka the
simulator security model- and hyperproperty-style formulations of
program trace behavior \cite{8429300,wysstar} including probabilistic
noninterference. In this Section we show for to formulate extensional
properties relevant to MPC passive security in our language model by
direct appeal to program distributions. In particular,
\emph{noninterference modulo output ($\NIMO$)} is shown to be sound
for MPC passive security. In Section \ref{section-automation} we will
show how $\NIMO$ can be automatically enforced directly via
whole-program analysis, or by automatically verifying composable,
intensional properties of circuit components.

@@

The real/ideal model posits a \emph{simulator} that exists in an ideal
world, whose goal is to reconstruct observable information in a real
world run of a protocol $\prog$ that correctly implements an ideal
functionality $\idealf$- that is, execution of a configuration
$\config{\store}{\prog}$. We will assume that any protocol $\prog$
includes a public \emph{output view} $\outv$ where the result is
published, where $0$ is the ``public'' client and thus $\outv$
is of the form $\view{0}{w}$ for some $w$, i.e., is owned by 0.
Note that the protocol is \emph{correct} iff the resulting
memory contains $\outv \mapsto \idealf(\store_S)$, where $S$ are the
secret inputs (honest and corrupt) in $\prog$.

@@

\subsection{Probabilistic Noninterference}
\label{section-hyperprop-ni}
Previous work has noted
relations between probabilistic noninterference and security
properties such as memory trace obliviousness \cite{darais2019language} and passive
security \cite{barthe2019probabilistic}. And above (Section \ref{section-pmf-examples}) we
have demonstrated that our basic encryption scheme enjoys a noninterference
property. To formulate probabilistic noninterference in
our setting we begin by defining low equivalence of memories.
\begin{definition}[Memory Low Equivalence]
  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
\end{definition}
Now, we can define probabilistic noninterference as a property of
programs ensuring that low-equivalent initial memories produce
the same low-observable trace (i.e., corrupt views) \emph{with the
same probability}. 
\begin{definition}[Probabilistic Noninterference]
  \label{definition-PNI}
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
  $\prog$ satisfies \emph{probabilistic noninterference}, and we write
  $\pni(\prog)$, iff for all
  $\store_1, \store_2 \in \mems(S)$:
  $$\store_1 =_C \store_2 \implies
    \condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2}$$
\end{definition}

Despite the intuitive resonance between noninterference
and passive security, since MPC
protocols implement an ideal functionality, they cannot generally satisfy
noninterference. We can easily demonstrate this by using
Example \ref{example-gmw-andcircuit} as a failure witness, where we'll
assume $C = \{0,2\}$ and $H = \{1\}$:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 0, \sx{2}{s2} \mapsto 1 \} \})
= 0
$$
whereas:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 1, \sx{2}{s2} \mapsto 1 \} \})
= 1
$$

@@

\subsection{A Sound Hyperproperty for Passive Security}
\label{section-nimo}
The intuition that passive security means that the protocol releases
no more information than what is released by conditioning on the
public output and corrupt inputs suggests a
refinement of noninterference. That is, the set of low equivalent
initial memories \emph{that can produce the same output} should generate the
same distributions of adversarial views.
\begin{definition}[Noninterference Modulo Output]
  \label{definition-NIMO}
  We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
  written $\NIMO(\prog)$, iff for all $H$ and $C$ with $|C|\le|H|$ and 
  $\store_1,\store_2 \in \mems(S)$ we have:
  $$
  (\store_1 =_C \store_2 \ \wedge \ 
  (\condd{\progd(\prog)}{\{ \outv \}}{\store_1} = \condd{\progd(\prog)}{\{ \outv \}}{\store_2}))
  \implies 
  (\condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2})
  $$
  where $\iov(\prog) = (S,V)$.
\end{definition}

>>>>

As discussed in Section \ref{section-related-work} previous
authors have noted connections between the traditional formulation of
the MPC security model- aka the simulator security model- and
hyperproperty-style formulations of program trace behavior
\cite{8429300,wysstar} including probabilistic noninterference. In
this Section we show for to formulate extensional properties relevant
to MPC passive security in our language model by direct appeal to
program distributions. In particular, \emph{noninterference modulo
output ($\NIMO$)} is shown to be sound for MPC passive security. In
Section \ref{section-automation} we will show how $\NIMO$ can be
automatically enforced directly via whole-program analysis, or by
automatically verifying composable, intensional properties of circuit
components.

@@

The real/ideal model posits a \emph{simulator} that exists in an ideal
world, whose goal is to reconstruct observable information in a real
world run of a protocol $\prog$ that correctly implements an ideal
functionality $\idealf$- that is, execution of a configuration
$\config{\store}{\prog}$. We will assume that any $\idealf$ is
deterministic- while MPC generally allows for probabilistic ideal
functionalities, this restriction captures a large and important class
of protocols. We will assume that any protocol $\prog$ includes a
public \emph{output view} $\outv$ where the result is published, where
$0$ is the ``public'' client and thus $\outv$ is of the form
$\view{0}{w}$ for some $w$, i.e., is owned by 0.  Note that the
protocol is \emph{correct} iff the resulting memory contains $\outv
\mapsto \idealf(\store_S)$, where $S$ are the secret inputs (honest
and corrupt) in $\prog$. Note that this implies that a correct
protocol publicizes the same output on the same secret inputs given
any choice of random tape.

@@

\subsection{Probabilistic Noninterference ($\pni$)} 
\label{section-hyperprop-ni}
Previous work has noted
relations between probabilistic noninterference and security
properties such as memory trace obliviousness \cite{darais2019language} and passive
security \cite{barthe2019probabilistic}. And above (Section \ref{section-pmf-examples}) we
have demonstrated that our basic encryption scheme enjoys a noninterference
property. To formulate probabilistic noninterference in
our setting we begin by defining low equivalence of memories.
\begin{definition}[Memory Low Equivalence]
  Given $\store^1$ and $\store^2$ with $\dom(\store^1) = \dom(\store^2) = X$,
  we write $\store^1 =_C \store^2$ iff $\store^1_{X_C} = \store^2_{X_C}$.
\end{definition}
Now, we can define probabilistic noninterference as a property of
programs ensuring that low-equivalent initial memories produce
the same low-observable trace (i.e., corrupt views) \emph{with the
same probability}. 
\begin{definition}[Probabilistic Noninterference]
  \label{definition-PNI}
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, we say that
  $\prog$ satisfies \emph{probabilistic noninterference}, and we write
  $\pni(\prog)$, iff for all
  $\store_1, \store_2 \in \mems(S)$:
  $$\store_1 =_C \store_2 \implies
    \condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2}$$
\end{definition}

Despite the intuitive resonance between noninterference and passive
security, since MPC protocols implement an ideal functionality, they
cannot generally satisfy noninterference. Noting that the public
output is always a corrupt view, we can easily demonstrate this by
using Example \ref{example-gmw-andcircuit} as a failure witness, where
we'll assume $C = \{0,2\}$ and $H = \{1\}$:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 0, \sx{2}{s2} \mapsto 1 \} \})
= 0
$$
whereas:
$$
(\progd(\prog_{\ref{example-gmw-andcircuit}}))({\outv \mapsto 1}|\{\{\sx{1}{s1} \mapsto 1, \sx{2}{s2} \mapsto 1 \} \})
= 1
$$

@@

\subsection{Noninteference Modulo Output ($\NIMO$)}
\label{section-nimo}
The intuition that passive security means that the protocol releases
no more information than what is released by conditioning on the
public output and corrupt inputs suggests a
refinement of noninterference. That is, the set of low equivalent
initial memories \emph{that can produce the same output} should generate the
same distributions of adversarial views.
\begin{definition}[Noninterference Modulo Output]
  \label{definition-NIMO}
  We say that a program  $\prog$ satisfies \emph{noninterference modulo output},
  written $\NIMO(\prog)$, iff for all $H$ and $C$ with $|C|\le|H|$ and 
  $\store_1,\store_2 \in \mems(S)$ we have:
  $$
  (\store_1 =_C \store_2 \ \wedge \ 
  (\condd{\progd(\prog)}{\{ \outv \}}{\store_1} = \condd{\progd(\prog)}{\{ \outv \}}{\store_2}))
  \implies 
  (\condd{\progd(\prog)}{V_C}{\store_1} = \condd{\progd(\prog)}{V_C}{\store_2})
  $$
  where $\iov(\prog) = (S,V)$.
\end{definition}
